---
title: "Lab 9 Notes: Chi-Squared"
author: "German Silva"
date: "11/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Attach packages

library(tidyverse)
library(here)
library(janitor)
library(broom)

```


## Read in the data

```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

getting counts of lizards by species

```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```

## Make a subset

- UTST lizards
- Sites: IBPE (enclousure grassland), RABB (rabbit mesquiresite), "EAST" (east tarbush site)
- Tail condition is recorded (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail != "NA")
```

## Find counts and proportions of lizards with broken or whole tails at each site

```{r}
utst %>% 
  count(site, tail)
# this looks fine but it doesn't look like a true constingency table
```

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail) # much better for chi-squared and communication but not for data analysis


tail_proportions <- tail_counts %>% 
  janitor::adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% 
  janitor::adorn_ns()

```


## Convert column to row names to create contingency table

```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Chi Squared

- Are tail breakage and site associated (independent)?
- Are these proportions so different do we have enough evidence to reject the H~o~ that there these variables are independent?

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

# there is a significant association between site and tail breakage

lizard_tidy <- broom::tidy(lizard_x2)

```

There is a significant association (i.e. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`)= `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`).



















